Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-10-09 2016-Weak-FH-Bisim/WeakBisim.tex
work with rabea ; sec 5 ; proofs

\documentclass{lncs/llncs}

@@

%\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{example}[theorem]{Example}
%\newtheorem{problem}[theorem]{Problem}
%\newtheorem{proof}[theorem]{Proof}

\title{Compositional equivalences based on Open pNets\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{Ludovic Henrio\inst{1}  \ \ \  Eric Madelaine\inst{1,2} \ \ \ Rab\'ea}
\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
}
\date{}                                           % Activate to display a given date or no date

@@

@@

\paragraph{Structure}
This article is organised as follows. Section~\ref{sec:notations}
provides the definition of pNets and introduces the notations used in
this paper. Section~\ref{sec:OT} defines open automata, i.e. automata
with parameters and transitions conditioned by the behaviour of
``holes''; a strong bisimulation equivalence for open automata is also
presented in this section. Section~\ref{section:op-semantics} gives
the semantics of open pNets expressed as weak automata, and states
compositional properties on the strong bisimulation for open
pNets. Finally, Section~\ref{sec:weak} defines a weak bisimulation
equivalence on open automata and derives weak bisimilarity for pNets,
together with properties on compositionality of weak bisimulaiton for
open pNets. Finally, Section~\ref{section:conclusion} concludes the
paper.

\section{Related works}

\TODO{TODO ... }

@@

pNets are composition entities equipped with a rich synchronisation mechanism: synchronisation vectors allows the expression of synchronisation between any number of entities and at the same time the passing of data between processes. Their strongest feature is that the data emitted by processes can be used can be used inside the synchronisation vector to do addressing: it is easy to synchronise a process indexed by $n$ with the action $a(v,n)$ of another process. This is very convenient to model systems and encode futures or message routing. pNets have been used to model GCM distributed component systems, illustrating the expressiveness of the model~\cite{AmeurBoulifa2017}. These works show that pNets are convenient to express the behaviour of the system in a compositional way, which is crucial for the definition of the semantics, especially when dealing with a hierarchical component system like GCM. Unfortunately, the semantics of pNets and the existing tools at this point were only able to deal with a closed system completely instantiated: pNets could be used as composition operator in the definition of the semantics, which was sufficient to perform finite-state model checking on a closed system, but there was no theory for the use of pNets as operators and no tool for proving properties on open system. The semmantics of closed pNets~\cite{AmeurBoulifa2017}, defined as an instantiation of labelled transition systems, is not necessary to understand this article but can be useful to  understand the semantics of pNets in a simpler and more operational setting. The theory of pNets as operators able to fully take into account open systems is given in the rest of this article.

@@

\section{Open Automata: A model of process composition}
\label{sec:OT}

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications. This section defines open automata and a bisimulation theory for them. Open automata are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on variables (potentially with only symbolic values).
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $\beta_j$, $s$, $\alpha$. \Post\ is a set of 
	equations 
	that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are variables of $s'$, and $e_k$ are expressions over the other 
	variables of the open transition. Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}

@@

\subsection{Bisimulation for open pNets and Composability}
\label{section:bisimulation-PN}
As  our symbolic operational semantics provides an open automaton, we can apply the notion of
	strong (symbolic) bisimulation on automata to open pNets:
\begin{definition}[FH-bisimulation for open pNets]\label{def:bisim-pnets}
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
\end{definition}
We can now prove that pNet composition  preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a pNet is filled by two bisimilar other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation (Theorem~\ref{thm-equiv}).

@@

Weak symbolic bisimulation was introduced to relate transition systems
that have indistinguishable behaviour, with respect to some definition
of \emph{internal actions} that are considered local to some
subsystem, and consequently cannot be observed, nor used for
synchronisation with their context.
As the notions of invisible action varies in different contexts,
e.g. $tau$ in CCS, and $i$ in Lotos, we will rely here on some set of
\emph{internal actions} depending on a specific action
algebra. Naturally, an internal action cannot be synchronised with
actions of other systems in its environment. We show here that under such assumption of non-observability of internal actions, we can define a weak bisimulation relation that is compositional, in the sense of open pNet composition. In this section we will first define a notion of weak transition similar to open transition. In fact a weak open transition is made of several open transitions labelled as internal transitions, plus potentially one non-internal open transition. This allows us to define weak open automata, and a weak bisimulation relation based on these weak open automata. Finally, we apply this weak bisimulation to open pNets, obtain a weak bisimilarity relationship for open pNets, and prove that this relation has compositional properties.

@@

Intuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
We define $\Id(s)$ as the identity function on variables of state s.
\begin{definition}[Non-observability of silent actions]\label{def:Non-ObsTau}
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $A$, $\alpha$,  $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           \beta_j^{j\in J},\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] If there is a $j$ such that $\beta_j=\tau$ then we have \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s) \land J=\{j\})
\]
\end{enumerate}
\end{definition}
The first statement of the definition state that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a silent action from a hole. The condition $J=\{j\}$ is a bit restrictive, it could safely be replaced by $\forall j\in J.\, \beta_j=\tau$, allowing the other holes to perform silent transitions too (because these silent actions cannot be observed).

We then can add any number of $\tau$ transitions of the holes before or after any open transition freely. This property justifies the fact that we can abstract away $\tau$ transitions from holes in the definition of a weak open transition.

Classically, we start by defining a derived transition relation based
on sequences of invisible actions:

@@

@@

@@

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a;\tau^*$ sequences from the OA, and build a weak
OT from each sequence. Here is one example of a short sequence,
starting with the open transition $OT_2$ of page \pageref{OT:ABP},
then transmitting the message to the receiver (if the sender and
receiver control bits are equal), and to the environment. The two
firts OTs are internal, the last one is visible, so it constitute a
$\tau;\tau;a$ weak transition.

@@

\section{Related Works}

@@

       	This is verified for each $p\in P$. Overall,  we have a family of open 
       	transitions 
       	$OT_{pxy}^{p\in 
       		P, x\in X, 
       		y\in Y} \subseteq \mathcal{T}_3$ that should simulate \emph{OT}.

@@

       \section{Proofs of properties on pNets}
\subsection{Composition Lemmas}       
       The proof of the other theorems rely on two main lemmas,
dealing respectively with the decomposition of a composed behaviour
between the context and the internal pNet, and with their recomposition. 
We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:

\begin{lemma}[OT decomposition]\label{lem-decompose} Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

Now focus on OTs of the subnets. For each $m\in I_1\uplus I_2$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\
for $m$ in $I_1$
\[
\pNet_m \models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
Or, $m$ in $I_2$
\[{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}}}\]

@@

Lemma \ref{lem-compose} is combining an open transition of $P$ with
an open transition of $Q$, and building a corresponding transition of
$P[Q]_{j_0}$, assembling their predicates.

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\TODO{important comment added:}
Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is not 
be satisfiable if the action of $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is now only used as an intermediate term inside formulas: it 
does not appear neither as global action nor as an action of a hole.

@@

\TODO{attention a l ordre: inverse d avant mais colle mieux a la notation -- check everywhere below}
\subsection{Weak bisimulation is an equivalence}

@@

 and similarly for $\Pred_{4i}$, $\Pred_{5i}$, $\Post_{4i}$, and $\Post_{5i}$  \TODO{define or OK this way?}

>>>>

\documentclass[preprint]{elsarticle}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{prop}{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{example}{Example}
%\newtheorem{problem}[theorem]{Problem}
\newenvironment{proof}{\paragraph{Proof}}{}

\title{Compositional equivalences based on Open pNets\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{Ludovic Henrio  \ \ \  Eric Madelaine \ \ \ Rab\'ea}
%\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
%	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
%}
\date{}                                           % Activate to display a given date or no date

@@

\tableofcontents

@@

\paragraph{Structure}
This article is organised as follows. Section~\ref{sec:notations}
provides the definition of pNets and introduces the notations used in
this paper, including the definition of open pNets. 
Section~\ref{sec:OT} defines open automata, i.e. automata
with parameters and transitions conditioned by the behaviour of
``holes''; a strong bisimulation equivalence for open automata is also
presented in this section. 
Section~\ref{section:op-semantics} gives
the semantics of open pNets expressed as open automata, and states
compositional properties on the strong bisimulation for open
pNets. 
Section~\ref{sec:weak} defines a weak bisimulation
equivalence on open automata and derives weak bisimilarity for pNets,
together with properties on compositionality of weak bisimulaiton for
open pNets. 
Finally, Section~\ref{sec:RW} discusses related works and Section~\ref{section:conclusion} concludes the
paper.

@@

pNets are composition entities equipped with a rich synchronisation mechanism: synchronisation vectors allows the expression of synchronisation between any number of entities and at the same time the passing of data between processes. Their strongest feature is that the data emitted by processes can be used can be used inside the synchronisation vector to do addressing: it is easy to synchronise a process indexed by $n$ with the action $a(v,n)$ of another process. This is very convenient to model systems and encode futures or message routing. pNets have been used to model GCM distributed component systems, illustrating the expressiveness of the model~\cite{AmeurBoulifa2017}. These works show that pNets are convenient to express the behaviour of the system in a compositional way, which is crucial for the definition of the semantics, especially when dealing with a hierarchical component system like GCM. Unfortunately, the semantics of pNets and the existing tools at this point were only able to deal with a closed system completely instantiated: pNets could be used as composition operator in the definition of the semantics, which was sufficient to perform finite-state model checking on a closed system, but there was no theory for the use of pNets as operators and no tool for proving properties on open system. The semantics of closed pNets~\cite{AmeurBoulifa2017}, defined as an instantiation of labelled transition systems, is not necessary to understand this article but can be useful to  understand the semantics of pNets in a simpler and more operational setting. The theory of pNets as operators able to fully take into account open systems is given in the following sections.

@@

\section{A model of process composition}
\label{sec:OT}

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications. This section defines open automata and a bisimulation theory for them.

\subsection{Open Automata}
 Open automata are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on variables (potentially with only symbolic values).
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $\beta_j$, $s$, $\alpha$. \Post\ is a set of 
	equations 
	that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are variables of $s'$, and $e_k$ are expressions over the other 
	variables of the open transition. Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}

@@

\subsection{pNet Composition Properties: composition of open transitions}
The semantics of open pNets allows us to prove two crucial properties relating pNet composition with pNet semantics: open transition of a composed pNet can be decomposed into open transitions of its composing sub-pNets, and conversely, from the open transitions of sub-pNets,  an open transition of the composed pNet can be built.

We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:
\begin{lemma}[Open transition decomposition]\label{lem-decompose} Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

Lemma \ref{lem-compose} is combining an open transition of $P$ with
an open transition of $Q$, and building a corresponding transition of
$P[Q]_{j_0}$, assembling their predicates.
\paragraph{Lemma~\ref{lem-compose}} {\bf Open transition composition}\\
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]

Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is often not  satisfiable, for example if the action  $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is  only used as an intermediate term inside formulas in the composed open transition: it 
does not appear neither as global action nor as an action of a hole.

\subsection{Bisimulation for open pNets -- a Composable Bisimulation Theory}
\label{section:bisimulation-PN}
As  our symbolic operational semantics provides an open automaton, we can apply the notion of
	strong (symbolic) bisimulation on automata to open pNets:
\begin{definition}[FH-bisimulation for open pNets]\label{def:bisim-pnets}
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
\end{definition}
We can now prove that pNet composition  preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a pNet is filled by two bisimilar other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation (Theorem~\ref{thm-equiv}).

@@

Weak symbolic bisimulation was introduced to relate transition systems
that have indistinguishable behaviour, with respect to some definition
of \emph{internal actions} that are considered local to some
subsystem, and consequently cannot be observed, nor used for
synchronisation with their context.
The notions of non-observable action varies in different contexts,
e.g. $tau$ in CCS, and $i$ in Lotos, we could define classically a set of
\emph{internal/non-observable actions} depending on a specific action
algebra. In this paper, to simplify the notations, we will simply use $\tau$ as the single non-observable action; the generalisation of our results to a set of non-observable actions is trivial. 
Naturally, a non-observable action cannot be synchronised with
actions of other systems in its environment. 
We show here that under such assumption of non-observability of $\tau$ actions, see Definition~\ref{def:Non-ObsTau}, we can define a weak bisimulation relation that is compositional, in the sense of open pNet composition. In this section we will first define a notion of weak open transition similar to open transition. In fact a weak open transition is made of several open transitions labelled as non-observable transitions, plus potentially one observable open transition. This allows us to define weak open automata, and a weak bisimulation relation based on these weak open automata. Finally, we apply this weak bisimulation to open pNets, obtain a weak bisimilarity relationship for open pNets, and prove that this relation has compositional properties.

@@

We first specify in terms of open transition, what it means for an action to be non-observable. Namely, we constraint ourselves to system where the emission of a $\tau$ action by a sub-pNet cannot be observed by the surrounding pNets. In other words, a pNet cannot change its state, or emit a specific observable action when one of its holes emits a $\tau$ action.

More precisely, we state that $\tau$ is not observable if the automaton always allows any $\tau$ transition from holes, and additionally the global transition resulting from a $\tau$ action of a hole is a $\tau$ transition not changing the pNet's state.
We define $\Id(s)$ as the identity function on variables of state s.
\begin{definition}[Non-observability of $\tau$ actions]\label{def:Non-ObsTau}
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe $\tau$ actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have:
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\beta_j$, $J$,  $\alpha$,  $s$, $s'$, $\Pred$, $\Post$  such that
\[ \openrule
         {
           \beta_j^{j\in J},\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] If there exists $j$ such that $\beta_j=\tau$ then we have: \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s) \land J=\{j\}
\]
\end{enumerate}
\end{definition}
The first statement of the definition states that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a $\tau$ action from a hole. The condition $J=\{j\}$ is a bit restrictive, it could safely be replaced by $\forall j\in J.\, \beta_j=\tau$, allowing the other holes to perform $\tau$ transitions too (because these $\tau$ actions cannot be observed).

Consequently, it is possible to add any number of $\tau$ transitions of the holes before or after any open transition freely. This property justifies the fact that we can abstract away $\tau$ transitions from holes in the definition of a weak open transition.

\TODO{stopped here}

Classically, we start by defining a derived transition relation based
on sequences of invisible actions.

@@

\subsection{Composition properties: composition of weak open transitions}
Now the semantics of an open pNet is defined both as an open automaton, and as a weak open automaton. Like the open automaton, the weak open automaton features valuable composition properties. We can exhibit  a composition property and a decomposition property that relate open pNet composition with their semantics, defined as weak open automata. These are technically more complex than the ones for open automata because each hole performs a set of actions, and thus a composed transition is the composition of one transition of the top-level pNet and a sequence of transitions of the sub-pnet that fills its hole.

@@

\subsection{Properties of weak bisimulation for open pNets}

@@

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a;\tau^*$ sequences from the OA, and build a weak
open trnsition from each sequence. Here is one example of a short sequence,
starting with the open transition $OT_2$ of page \pageref{OT:ABP},
then transmitting the message to the receiver (if the sender and
receiver control bits are equal), and to the environment. The two
firsts open transitionTs are internal, the last one is visible, so it constitute a
$\tau;\tau;a$ weak transition.

@@

\section{Related Works}\label{sec:RW}

@@

       	This is verified for each $p\in P$. Overall,  we have a family of open 
       	transitions 
       	$OT_{pxy}^{p\in 
       		P, x\in X, 
       		y\in Y} \subseteq \mathcal{T}_3$ that should simulate $OT$.

@@

       \section{Proofs of properties on pNets}
\subsection{Composition Lemmas}       
       The proof of the other theorems rely on two main lemmas,
dealing respectively with the decomposition of a composed behaviour
between the context and the internal pNet, and with their recomposition.

\paragraph{Lemma~\ref{lem-decompose}}: {\bf Open transition decomposition} \\
 Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.

@@

Now focus on OTs\footnote{In the appendix we use OT to abbreviate ``open transition'' and WOT to stand for ``weak open transition.''} of the subnets. For each $m\in I_1\uplus I_2$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\
for $m$ in $I_1$
\[
\pNet_m \models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
Or, $m$ in $I_2$
\[{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}}}\]

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}

Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is not 
be satisfiable if the action of $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is now only used as an intermediate term inside formulas: it 
does not appear neither as global action nor as an action of a hole.

@@

\subsection{Weak bisimulation is an equivalence}

In this section, we first define two alternative definitions for weak bisimulation. We use these two alternative definitions to show that weak bisimulation is an equivalence, we will also re-use these alternative definitions in the  proofs of the theorems in next sections.

@@

 and similarly for $\Pred_{4i}$, $\Pred_{5i}$, $\Post_{4i}$, and $\Post_{5i}$.
